---
title: "Lab 5"
author: "Ashlan Simpson"
date: "4/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

```


1) Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?
```{r}
library(lubridate)
acc$Weekday <- wday(acc$DAY_WEEK, label = T)
counts <- table(acc$Weekday)
barplot(counts)
```
2) Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)
```{r}
Fatal <- filter(per, INJ_SEV==4)
```
3) Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)
```{r}
A <- Fatal %>% group_by(STATE) %>%
  count(MAKE) 
A <- A[complete.cases(A), ]

mostDangerous <- A %>%
  group_by(STATE) %>% 
  filter(n == max(n))

```
4) Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)
```{r}
library(ggplot2)
library(tidyverse)
Q4 <- readxl::read_xlsx("DF.xlsx")
mostDangerous$STATE <- as.character(mostDangerous$STATE)
Q42 <- inner_join(mostDangerous, Q4, by=c("STATE"="State Code"))
states <- map_data("state")
states$region <- toupper(states$region)
Q43 <- inner_join(Q42, states, by=c("State Name"= "region"))



ggplot(states, aes(x=long, y=lat)) + geom_polygon(aes(group=group)) + geom_text(aes(label = Make))
```



5) Join the accident and person table (work out which variable(s) to use)
```{r}
Full <- left_join(per, acc)
```
6) Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX).
Visualize the results and explain what you find.
```{r}
library(ggrepel)
Data2 <- All %>% group_by(HOUR, Weekday, Sex) %>%
  summarise(Deaths = n()) %>%
  filter(HOUR == 23)

All %>% group_by(HOUR, Weekday, Sex) %>%
  summarise(Deaths = n()) %>%
  ggplot(aes(x=HOUR, y=Deaths, group=Weekday)) + geom_line() + facet_wrap(~Sex) + geom_text_repel(data=Data2, aes(x=HOUR, y=Deaths, label=Weekday))
```






