```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
per <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```


1) Are there some days of the week where more accidents happen than the others (see FARS manual, use variable DAY_WEEK)?
```{r}
library(lubridate)
acc$Weekday <- wday(acc$DAY_WEEK, label = T)
counts <- table(acc$Weekday)
barplot(counts)
```
People were hurt more on Saturdays
2) Create a data frame containing the persons who are fatally hurt in the accidents (look up variable INJ_SEV)
```{r}
Fatal <- filter(per, per$INJ_SEV==4)
```
3) Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)
```{r}
library(dplyr)
library(tidyverse)
A <- Fatal %>% group_by(STATE) %>%
  count(acc$MAKE) 
A <- A[complete.cases(A), ]
mostDangerous <- A %>%
  group_by(STATE) %>% 
  filter(n == max(n))
```
4) Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)
```{r}
library(maps)
library(ggplot2)
library(tidyverse)
Q4 <- readxl::read_xlsx("DF.xlsx", guess_max = min(2000, n_max = NULL))
mostDangerous$STATE <- as.character(mostDangerous$STATE)
Q42 <- inner_join(mostDangerous, Q4, by=c("STATE"="State Code"))
states <- map_data("state")
states$region <- toupper(states$region)
holder<- states
```

```{r}
library(usmap)
StateLonLat <- mostDangerous %>%
  group_by(STATE) %>%
  right_join(states, by = c(STATE = "region")) %>%
  select(STATE, long, lat)
StateLL <- aggregate(cbind(StateLonLat$long,StateLonLat$lat), by = list(STATE = StateLonLat$STATE), FUN = mean)
StateLL <- StateLL %>% 
  rename(
    long = V1,
    lat = V2
)
danger <- aggregate(Q42$MAKE, by = list(STATE = Q42$"State Name"), FUN = mean)
fin <- merge(danger, StateLL, by ="STATE")
fin<- fin %>%
  rename(
    Make = x
  )
states <- aggregate(states$group, by = list(STATE = states$region),FUN = mean)
states<- states %>%
  rename(
    group = x
  )
fin <- merge(fin,states, by = "STATE")
```
```{r}
mainStates<-map_data("state")
ggplot() + 
  geom_polygon( data=mainStates, aes(x=long, y=lat, group=group),
                color="black", fill="lightblue" ) + geom_text(data = fin, aes(x=long,y=lat, label = Make))
```



5) Join the accident and person table (work out which variable(s) to use)
```{r}
Full <- left_join(per, acc)
```
6) Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX).
Visualize the results and explain what you find.
```{r}
library(ggrepel)
Data2 <- Full %>% group_by(HOUR, Weekday, SEX) %>%
  summarise(Deaths = n()) %>%
  filter(HOUR == 23)
Full %>% group_by(HOUR, Weekday, SEX) %>%
  summarise(Deaths = n()) %>%
  ggplot(aes(x=HOUR, y=Deaths, group=Weekday)) + geom_line() + facet_wrap(~SEX) + geom_text_repel(data=Data2, aes(x=HOUR, y=Deaths, label=Weekday))
```
This graph (Men on the right Women on the left) shows that death rates for men and women both spike on the weekend but men Already have a much higher deat rate. In hours of the day both see a slight rise in the morning this could be due to a morning commute or workplace fatalities.




7) Now plot a choropleth map of the number of deaths on a county level. Also explain what you find.
```{r}
temp <- Q4 %>%
  rename(
    County = "County Name"
  )
county <- aggregate(cbind(counties$long, StateLonLat$lat), by = list(County = counties$subregion), FUN = mean)
county <- county %>%
  rename(
    long = V1,
    lat = V2,
  )
county$County <- tolower(county$County)
temp$County <- tolower(temp$County)
here <- merge(temp,county,by = "County")
fatals <-fatalities
fatals <- fatalities %>%
  rename(
    CCode = "COUNTY",
    SCode = "STATE"
  )
temp <- Q4 %>%
  rename(
    CCode = "County Code",
    SCode = "State Code"
  )
#countyfin<-county%>%
 # rename(
  ## SCode = "State Code"
  #)
county
total <- merge(temp,fatals,by=c("CCode","SCode")) 
total$"County Name" <- tolower(total$"County Name")
total1 <- total %>%
  rename(
    County = "County Name"
  )
total1
last<-merge(total1,county,by= "County")
last
counties
county1 <- aggregate(counties$group, by = list(county = counties$subregion),FUN = mean) 
county1
```
```{r}
mainCounties<-map_data("county")
ggplot(last, aes(x=long, y= lat, fill = nFatal)) + geom_polygon( aes(x= long, y=lat))
```
```
8) Is summer or winter more dangerous? Does this depend on states? Explore and explain.
```{r}
#When output actually works replace this text with the answer to the question
ggplot(Full, aes(x=LONGITUD, y=LATITUDE, fill=WEATHER)) + geom_polygon()
```